Framing Effects in the Wild

### 2022-08-22 paper/Manuscript_Framing.Rmd
Update abstract, add change across time

---
title: "Framing Effects in the Wild"
author: "Nicolai Berk"
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    toc: false
    fig_caption: yes
    includes:
      in_header: preamble.tex
# subtitle: Manuscript prepared for the 12th Annual Conference of the European Political Science Association in Prague, Czech Republic
fontsize: 12pt
geometry: margin=1.5in
abstract: "A large body of evidence from experimental work suggests that individual attitudes can easily be altered by emphasizing different aspect of an the issue at stake. However, such experimental studies of *emphasis framing* are limited in their emulation of the real world, where consumers are treated by multiple sources, select their news content, and pay more or less attention compared to the experimental setting. I exploit a rare change in news framing in the major German tabloid paper - *Bild* - to study whether experimental estimates of framing effects translate to the real world. In a first step, I apply supervised transformer models to a novel dataset of 2.5M news articles and provide evidence of a substantial shift in *Bild*'s migration framing following an editorial change. Second, I use a difference-in-differences design on a panel datset of over 16,000 individuals in ten waves to assess the precise impact on migration attitudes. The estimated effect is thirty times smaller than similar estimates from experimental studies, statisticially indistinguishable from zero, and significantly smaller than even minimal relevant effect magnitudes. Further analyses show that no change can be observed on other, related dependent variables either, and suggest that the change did not result in outward or inward selection into the newspaper. Taken together, these findings fundamentally challenge the external validity of framing effects research and suggests more limited influence of the news media on public opinion than previously thought."
bibliography: C:/Users/nicol/Dropbox/Studium/BibTex/PhD.bib
---

@@

WORD COUNT:

*I am thankful for extensive feedback and other support on this manuscript from Christian Arnold, Daniel Bischof, Philipp Broniecki, Jóhanna Ýr Bjarnadóttir, Elias Dinas, Florian Foos, Felix Hagemann, Violetta Haas, Julian Heid, Christina Gahn, Heike Klüver, Martin Vinæs Larsen, Matthew Levendusky, Will Lowe, Thomas M. Meyer, Rune Slothuus, Judith Spirig, Antonio Valentim as well as participants in the EPSIP Colloquium at HU Berlin, CIVICA EGN Workshop at the EUI, CompText Conference 2022, and the Political Behaviour Section at Aarhus University. I also thank the two RAs, Robin Groß and Nathalie Aerts, for annotating the news articles. All remaining confusions are my own.*

@@

The persuasiveness of the news media has been a fixation of research on political communication and public opinion for at least the past 100 years. One of the largest fields within this literature is concerned with framing effects. A huge body of experimental evidence suggests that citizens substantially alter their issue attitudes when presented with different frames [@Busby2019; @Chong2007b; @Leeper2020; @Nelson1996; @Nelson1997Rally; @sniderman2004structure; @Slothuus2008a]. However, there is little research investigating whether these findings translate into a messy reality [@Barabas2010], where frames are communicated by specific actors, and citizens self-select into media outlets, potentially shielding them from some frames and exposing them to others.

In fact, many researchers argue that the effects of media content on attitudes are minimal. Observational research on media effects is abundant [@Broockman2022; @Durante2012; @Foos2022; @Grossman2022; @Guess2021; @Levendusky2022; @Martin2017; @Spirig2020], but has rarely paid attention to the precise content affecting issue attitudes [see @Djourelova2020 for an exception]. To the authors knowledge, no existing observational study assesses the causal effect of changing news framing on issue attitudes.

I address this gap in the literature by presenting evidence from an editorial change in the largest German tabloid *Bild*, resulting in an increasing emphasis of crime in the paper's migration coverage. The case is ideal to assess the everyday influence of issue frames on public opinion, as one outlet changed it's framing while other papers maintained theirs. By employing BERT models for the measurement of crime frames in a corpus of 2.5 million newspaper articles, I am able to clearly identify the treatment - something that is usually left to experimental studies. Coupling this data with a large-scale panel survey of over 15,000 individuals allows the precise estimation of this framing effect with a difference-in-differences design.

The results show that emphasis framing does not affect citizens' political attitudes in the real world. The influence of the news media on opinion formation in contemporary news environments with high levels of consumer choice and diverse exposure seems to be rather limited. Similarly, the study contributes to the media effects literature by investigating the effects of changing news *content* on public opinion, independent of the effects of the changing availability of news *outlets*. More generally, this research has important implications for our understanding of contemporary democracies, questioning the role of the media in public opinion formation.

The paper will proceed as follows: in the next section I will briefly discuss the existing literature on framing and media effects, before describing the theoretical framework and research design, including case, measurement, and estimation strategy. The results section will then showcase findings from difference-in-differences and fixed-effect models assessing the impact of increased emphasis of crime in news coverage about migration, both on the association of attitudes about crime with attitudes about migration, as well as migration attitudes themselves. Lastly, I discuss the wider implications of the findings for the study of framing effects and the role of the news media in the opinion formation process.

<!-- look at rune's comments regarding this chapter - keep focusing on framing effects literature -->

Research on framing has shown for decades that citizens' attitudes are highly reactive to minor changes in question wording and emphasis of specific considerations. However, it remains unclear how well these experimental findings translate into the real world, where citizens self-select into media content, receive cues from political actors, and are usually more concerned with their private lifes than politics. Scholarly effort has therefor been devoted to test the conditions under which framing effects materialize, designing experiments which emulate different aspects of real-world environments. For example,it has been shown that exposure to conflicting frames can serve to strengthen the effect of frames one was earlier exposed to, dependent on individual preconditions [@Bechtel2015; @Chong2013; @Druckman2012]. Others have assessed what influence the sender of a frame has and shown that the effects of framing are more pronounced when the sponsor of the frame is from one's own party [@Druckman2001; @Slothuus2010b; @Slothuus2010]. Lastly, concerns about the duration of framing effects have been addressed by some experimental research showing that those effects can last at least several weeks [@Lecheler2011]. (missing here: Lau et al)

While these findings enable a better understanding about individuals' susceptibility to framing effects in real-world environments, the literature almost exclusively provides us with experimental evidence. Experiments can only incorporate few characteristics at a time and are limited to emulate a realistic environment for the study of news framing. As Donald Kinder puts it: "[E]xperimental results can always be questioned on their generalizability, and framing effects are no exception. [...] A more balanced reading of frame effects requires methodological diversification, experiments and studies oriented to the world outside" [-@Kinder2007, 157]. In order to test the external validity of framing effect studies, it is necessary to study observational data.

This call for an assessment of framing effects beyond experimental studies has been echoed by various scholars over time [@Busby2019; @Leeper2020], yet rarely been answered. Existing research is often riddled with endogeneity concerns, co-treatment or concerned with media effects broadly, instead of assessing framing effects specifically. For example, @Jerit2008 investigates the effects of framing on public opinion compared to addressing the opponents issue. She finds that reframing is less effective than engagement with the opposite parties' frames, a finding that suggests that framing is not very effective. However, this study does not try to identify causal effects, but restricts itself to the assessment of correlational evidence. Others show with similar correlational designs that only specific frames seem to drive issue importance [@Mclaren2018], but do not assess effects on attitudes.

Other work has put more emphasis on causal identification of media effects using quasi-experimental designs, but paid less attention to the precise definition of the independent variable. This literature has usually exploited changes in news media markets such as boycotts of certain outlets [@Foos2022], entries of new outlets [@DellaVigna2007; @Grossman2022], changing TV channel positions [@Martin2017], or partisan takeovers of news outlets [@Durante2012; @Levendusky2022; @Spirig2020] to estimate the effects of changing news exposure on political behavior. Interestingly, the results diﬀer across studies. While some studies find substantial effects of exposure to slanted media on political attitudes and voting behavior even in the long run [@Foos2020; @Grossman2022], others find that consumers switch outlets when the slant of currently consumed news changes [@Durante2012; @Spirig2020].

Another approach is to assess consumer reactions to news content through ﬁeld experiments. In this setup, consumers are randomly assigned to speciﬁc news outlets and the development of their political attitudes is assessed. @Guess2021, for example, randomly assigned individuals to receive content from a left-leaning or a right-leaning news outlet in their online news feeds. Similar studies have assessed the ffects of cross-partisan content on TV [@Broockman2022] and social media [@Bail2018]. This literature provided contradicting results as well: while some found liberalizing effects of liberal content [@Broockman2022], others found no effects [@Guess2021], or even partisan backlash (i.e. individuals becoming even more conservative in response to liberal content; @Bail2018).

I argue this lack of consensus regarding the effects of media content is at least partly related to a lack of precision in the definition of the independent variable. The vast majority of studies in this literature observes or manipulates the consumed *outlet* , rather than the *content*. This is problematic for three main reasons. First, by assuming a stable 'slant' of specific news outlets, it remains unclear which content consumers actually react to (or not). While it is certainly true that news outlets have a certain political leaning, there is variation in how political and how persuasive specific news content is. Second, citizens are often aware of outlets' political leaning. @Baum2008a showed that consumers place the same news content more left- or right-leaning, dependent on wether it was presented as Fox news or CNN content. This implies that citizens should be able to discount political bias of outlets (muting potential effects of biased content) or might show motivated responses to content they perceive as motivated by political motives (e.g. backlash to opposing news sources). Third, by focusing on the effects of the consumption of different outlets, the literature studies a process unlikely to be observed in the real world - after all, changes in the makeup of media markets are extremely rare and switches from one partisan outlet to an opposing partisan outlet even more so. The main influence of news editors on public opinion should be exercised by editing the everyday content of news, rather than changing the consumer base of their outlets.

To my knowledge, there are only two observational studies assessing the causal effect of exposure to specific content on political attitudes. @Slothuus2010b takes advantage of a sudden shift in a political parties' framing of retirement benefits, and shows that this changed framing has substantial impact on party supporters' attitudes about the issue. However, it is unclear how these findings translate to news framing effects, as it is impossible to distinguish the effect of partisan cue-taking from the persuasive effect of the frame itself. The only study which assesses the precise impact of changing news content on political attitudes comes from @Djourelova2020. She exploits a ban of the term "illegal migrant" in the *Associated Press* and shows that this language has modest effects migration attitudes. While the study itself reveals impressively how language can shape political attitudes, it constitutes a highly specific case and a rather minor change in the framing of migration. While news stories might lack the term "illegal migrant", they might still extensively cover crimes committed by migrants and thus associate migration with crime.

Summarized, we lack an understanding of how framing affects political opinions in the real world. While experimental evidence suggests rather strong effects, it is unclear how these findings travel outside the lab. Observational evidence on media effects is abundant but has not studied precisely what the impact of framing is. The present study therefor exploits a naturally occurring change in the news framing of a particular issue *within* a specific outlet to isolate the effects of news framing on political opinions.

# Emphasis Framing and the Value-Expectancy Model

@@

Work using the emphasis framing concept of framing - as opposed to other definitions like equivalence frames [@Kahneman1979prospect], episodic frames [@Iyengar2005], or generic frames [@DeVreese2005] - mostly builds on the expectancy-value model developed by @Ajzen2000. This model suggests that an attitude on a given issue is a function of two things: considerations relating to that object and their respective weights. The evaluation of the issue is subdivided into a number of considerations, which will carry clear evaluations. For example, to form an opinion about whether immigration to one's country should be restricted, a person might consider the humanitarian conditions in the countries of origin (negative consideration regarding restrictive migration policy), as well as the risk of increased crime (positive consideration; note that the model does not differentiate whether a consideration is grounded in objective reality). Individuals weigh each of these considerations to form an overall opinion on the subject.

@@

```{=tex}
\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{vis/CognitiveStorage.png}
    \caption{The cognitive evaluation of an issue $y$ is determined by the strength of association $w_i$ with other concepts with existing evaluations $x_i$.}
    \label{fig:cogStor}
\end{figure}
```
According to this model, framing changes opinions by changing the association of a single consideration with the issue at hand. As a result, a specific consideration factors more prominently in an individuals' attitude formation process: "media frames influence opinions by stressing specific values, facts, or other considerations, endowing them with greater apparent relevance to the issue than they might appear to have under an alternative frame" [@Nelson1997Rally, 569].

@@

***Evaluation hypothesis*****: When an individual is exposed to a specific negative consideration about an issue, their attitude about the overall issue should become more negative.**

@@

In order to test whether changing news content - rather than a shift in the consumed news outlet - affected political attitudes, I exploit an editorial change in the largest German tabloid newspaper, *Bild*. As I will show, this editorial change resulted in a substantial shift in the papers' migration framing and is hence ideal to assess the effect of within-newspaper content shifts. Additionally, the fielding of a large-scale panel survey in the same period allow to assess the impact of news content on migration attitudes and how predispositions condition this effect. I will briefly discuss the context of the case in this section.

@@

In December 2016, Axel Springer, the company owning Bild, announces that Diekmann will leave the company. Less than a week after, it becomes public that prosecutors investigate Diekmann for sexual harassment. While the company claimed that this was not the reason for Diekmann's departure, the process might have been sped up as a result of the investigations\footnote{According to a former editor of the newspaper I talked to, rumors in the company at the time also suggested Diekmann's departure might have been sped up as a result of the investigation.} [@Spiegel2017Diekmann]. The migration coverage in *Bild* severely changed when Diekmann left the newspaper in February 2017 and Julian Reichelt took over the position as editor-in-chief. The paper increasingly started painting a picture of refugees as criminals on the tabloid's front page [@Niggemeyer2018Bild; @DLF2018Reichelt]. Headlines like "I killed 40 people and want asylum" increasingly dominated the migration coverage of the newspaper. In a speech months before he took over the position as chief editor, Reichelt outlined his motivation for such a shift in migration coverage: "I can assure you: nothing has hurt [*Bild*] economically as much as our clear, humane, empathetic stance in the refugee crisis" \footnote{Reichelt in a speech at Deutschlandfunk's conference "Formate des Politischen 2016": \url{https://vimeo.com/190347766}}. He also argued that the German media shows too little tolerance towards anti-immigration attitudes and therefor excludes those holding such views from the political discourse.

This presents a unique case to study the effects of media content on public opinion. As outlined above, past observational studies have usually explored the effect of consuming a new/different news outlet, without assessing the precise shift in content. In this case, the shift in content took place *within* the outlet and can be clearly identified. The change of editor itself was not salient, which means any opinion changes among readers are most likely the result of the content [@Spirig2020]. Additionally, the timing of the editorial change is potentially exogenous, given the pressure from allegations against Diekmann. In summary, *this case is as close as possible to a field experiment in which a specific consideration of migration coverage is amplified in the coverage of one newspaper, but not others*.

I estimate the effect of this communicative shift in the tabloid's migration coverage with a difference-in-differences design to answer the question how much influence the news media has to shape public opinion. Compared to the experimental setting, this can tell us with high external validity how news content affects political attitudes in contemporary news environments, in the medium term, in the real world.

## Measuring Crime Frames in Migration Coverage

To measure media attention to immigration, I collected 2.5 million newspaper articles from the websites of the most important German broadsheets \textit{Frankfurter Allgemeine Zeitung (FAZ), Spiegel Online (SPON), Süddeutsche Zeitung (SZ), Die Tageszeitung (TAZ), Die Welt} and the major German tabloid \textit{Bild} for the period 2013-2019. Following the value-expectancy model, I identify crime frames in migration coverage in a two-step process: first, I use supervised machine learning to identify content about migration. Second, I identify content about crime *within* migration coverage by applying a similar supervised classifier. This approach closely resembles the value-expectancy model, as I first identify coverage about the *issue*, before identifying the considerations emphasized in the discussion of this issue.

Table 1 in the appendix shows the number of print copies sold and online page impressions for each newspaper per day. It becomes clear that German newspapers have a rather important position in the German news market. The tabloid \textit{Bild} has by far the largest reach, with over 1.7 Million sold newspaper daily and around 550 million web visits per month. For reference: the German Wikipedia site had less than twice as many visits (975 million) in the same time frame \footnote{Wikipedia Siteviews Analyisis: https://pageviews.wmcloud.org/siteviews/}. *Bild* is closely followed by the weekly published \textit{Spiegel}. The daily broadsheets sell less, and interestingly the online presence seems to be dominated by the right, given the weak presence of centre-left \textit{SZ} and left-wing \textit{taz}.

To identify whether an article is about migration, a research assistant annotated 1800 articles. As migration content is rare, and training data needs to be somewhat balanced, articles cannot be selected randomly. Instead, I first construct a dictionary of terms related to migration. I use dictionary extension based on German GloVe word-embeddings\footnote{Downloadable from \url{https://deepset.ai/german-word-embeddings}.} to construct a comprehensive dictionary, and apply it to the articles. Based on the relative share of migration words in an article, I draw a stratified sample of 300 articles (100 from articles with no migration terms, 100 from the quarter with most migration terms and 100 from the remainder in between) for each newspaper for a total of 1,800 papers. Then, a coder hand-coded these articles, assessing whether their main topic is related to migration.

Using this sample, a BERT transformer model\footnote{\url{https://huggingface.co/bert-base-german-cased}} is fine-tuned on a subset of 1,400 annotated articles. After fine-tuning, the model correctly classifies 95.5% of the test set (F1: 0.94, recall: 0.93, precision: 0.95). This classifier is then used to annotate all 2.5 million newspaper articles. For the period from 2013 to 2019, around 90,000 (3.6%) are identified to treat the issue of migration.

@@

Figure \@ref(fig:salience) shows the share of migration articles in each newspaper. Clear peaks in attention are visible surrounding the refugee movements of summer 2015, following the large-scale sexual assaults on new years eve in Cologne 2016, and the conflict over asylum policy between the conservative sister parties CDU and CSU in 2018. Newspapers seem to be event-driven and follow largely parallel trends, albeit at different levels, with the centre-left \textit{SZ} discussing migration most and the tabloid \textit{Bild} least. Attention decreases in mid-2016, but remains high compared to pre-2015 levels. Surrounding the fielding period of the survey waves (gray vertical lines), most newspapers show similar levels of attention to the issue between 2.5% and 6% of all articles, with the exception of \textit{SZ}, which moves between 5% and 12.5%.

To identify migration coverage containing the crime consideration and thereby associating migration with crime, I first fine tune a BERT classifier for German language in a similar fashion as above, this time with coverage about crime as the outcome of interest. Using a training sample of 1400 articles and a test sample of 400 articles, the classifier reaches a very good performance with an F1 score of 85.7%. I then apply this classifier to classify the approximately 90,000 articles about migration identified by the first classifier. About 7% of all migration coverage consists of articles about some form of crime\footnote{See appendix 1.2 for more information about the classifier, training process and data.}. The share of migration content devoted to news about crime within a given time frame allows me to assess how prominent the consideration of crime as a result of migration was at a specific point in time.

>>>>

---
title: "Framing Effects in the Wild"
author: "Nicolai Berk"
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    toc: false
    fig_caption: yes
    includes:
      in_header: preamble.tex
# subtitle: Manuscript prepared for the 12th Annual Conference of the European Political Science Association in Prague, Czech Republic
fontsize: 12pt
geometry: margin=1.5in
abstract: "Does news framing affect public opinion? While a large body of experimental work suggests that individual attitudes can be easily altered by emphasizing different aspects of an issue, little is known about how these studies translate into the real world. I exploit a rare natural change of news framing in the major German tabloid newspaper *Bild* to study whether experimental estimates of framing effects translate into the real world. In a first step, I apply supervised transformer models to a novel dataset of 2.5M news articles and provide evidence of a substantial shift in the paper's migration framing following an editorial change. Second, I employ a difference-in-differences design using panel data from over 16,000 individuals in ten waves to assess the precise impact on migration attitudes. The findings show that, despite a major change in migration framing, immigration attitudes and related variables remain unaffected. News framing only seems to be effective within limited opportunity windows during the emergence of political issues."
bibliography: C:/Users/nicol/Dropbox/Studium/BibTex/PhD.bib
---

@@

WORD COUNT: `r wordcountaddin::word_count(here("paper/Manuscript_Framing.Rmd"))`

*I am thankful for extensive feedback and other support on this manuscript from Christian Arnold, Daniel Bischof, Philipp Broniecki, Jóhanna Ýr Bjarnadóttir, Elias Dinas, Florian Foos, Violetta Haas, Felix Hagemann, Julian Heid, Christina Gahn, Heike Klüver, Martin Vinæs Larsen, Matthew Levendusky, Will Lowe, Thomas M. Meyer, Rune Slothuus, Judith Spirig, Antonio Valentim as well as participants in the EPSIP Colloquium at HU Berlin, CIVICA EGN Workshop at the EUI, CompText Conference 2022, and the Political Behaviour Section at Aarhus University. All remaining confusions are my own.*

<!--# I also thank the two RAs, Robin Groß and Nathalie Aerts, for annotating the news articles.  -->

@@

Does news framing affect public opinion? While a large body of experimental evidence suggests that the emphasis of a certain consideration can substantially alter issue attitudes, we know little about how these findings translate into real-world environments. Prior work has shown that the external validity of such experiments might be seriously complicated by environments of diverse frames [@Chong2013] and actors [@Slothuus2010], as well as the seeking and avoidance of specific content [@Lau2021a; @Taber2006]. Contemporary news environments provide consumers with especially diverse content and maximize their abilities to choose specific outlets and avoid others [@Bennett2008], thereby increasing the gap between the real-world environment and experimental studies. Additionally, the very nature of experiments is prone to over-estimate effects in the real world [@Barabas2010].

Following similar considerations, Donald Kinder called for framing research to extend its toolkit: "[E]xperimental results can always be questioned on their generalizability, and framing effects are no exception. [...] A more balanced reading of frame effects requires methodological diversification, experiments and studies oriented to the world outside" [-@Kinder2007, 157; see also @Busby2019; @Leeper2020]. To the author's knowledge, no study has yet answered this call to provide an estimate of framing effects in the real world.

I address this gap in the literature by exploiting a rare and naturally occuring shift in the migration framing of Germany's largest tabloid newspaper, *Bild*. The case is ideal to assess the everyday influence of issue frames on public opinion, as one outlet substantially changed it's framing while other papers maintained theirs. By employing BERT models for the measurement of crime frames in a corpus of 2.5 million newspaper articles, I am able to clearly identify the treatment - something that is usually left to experimental studies. The fielding of the German Longitudinal Election Study in the same period provides the rare opportunity to precisely assess the effect of this changing framing on readers' migration attitudes in a *difference-in-differences (DiD) design*.

The results indicate that even this strong change in news framing did not affect consumers' immigration attitudes. The estimated effect is statisticially indistinguishable from zero, and significantly smaller than even minimal relevant effect magnitudes. Other variables, such as related issue attitudes, issue importance, and voting intentions are similarly unaffected. Additional evidence suggests that readers do also not react by turning away from the newspaper. Overall, these results fundamentally challenge the external validity of framing effects research and suggest that the influence of individual news outlets on public opinion is rather limited. Even strong changes in news framing do not affect political attitudes or news consumption.

The paper will proceed as follows: in the next section I will briefly introduce the theoretical framework of emphasis framing and derive hypotheses about the impact of news framing on political attitudes. I will then discuss the limited external validity of framing effects studies, before I turn to the research design, including case, measurement, and estimation strategy. The results section will then showcase evidence from difference-in-differences models assessing the impact of changing migration framing on migration attitudes. Lastly, I discuss the wider implications of the findings for the study of framing effects and our understanding of the role of news media in the opinion formation process.

# Emphasis Framing

@@

Work using the emphasis framing concept of framing - as opposed to other definitions like equivalence frames [@Kahneman1979prospect], episodic frames [@Iyengar2005], or generic frames [@DeVreese2005] - mostly builds on the expectancy-value model developed by @Ajzen2000. This model suggests that an attitude on a given issue is a function of two things: considerations relating to that object and their respective weights. The evaluation of the issue is subdivided into a number of considerations [a so-called "frame in thought"; @Chong2007a, 105; @Kinder1996], which will carry clear evaluations. For example, to form an opinion about whether immigration to one's country should be restricted, a person might consider the humanitarian conditions in the countries of origin (negative consideration regarding restrictive migration policy), as well as the risk of increased crime (positive consideration; note that the model does not differentiate whether a consideration is grounded in objective reality). Individuals weigh each of these considerations to form an overall opinion on the subject.

@@

```{=tex}
\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{vis/CognitiveStorage.png}
    \caption{The cognitive evaluation of an issue $y$ is determined by the strength of association $w_i$ with other concepts with existing evaluations $x_i$.}
    \label{fig:cogStor}
\end{figure}
```
How does this model incorporate news content and other forms of political communication? News content can emphasize different considerations of an issue [so called "frames in communication" @Kinder1996]. This strengthens the cognitive association of the issue at hand with the raised consideration. As a result, this specific consideration factors more prominently in an individuals' attitude formation process: "media frames influence opinions by stressing specific values, facts, or other considerations, endowing them with greater apparent relevance to the issue than they might appear to have under an alternative frame" [@Nelson1997Rally, 569].

@@

***Evaluation hypothesis*****: When an individual is exposed to a specific positive/negative consideration about an issue, their attitude about the overall issue should become more positive/negative.**

Research on framing has shown for decades that citizens' attitudes and their issue associations are highly reactive to minor changes in the emphasis of specific considerations. However, it remains unclear how well these experimental findings translate into the real world. Especially when it comes to the exposure to frames in communication, translating framing effects from experimental studies is only possible under strong assumptions: While framing studies usually simply present a news article or other forms of political content to respondents, news content in the real world competes with other news and entertainment content to reach its audience. Additionally, changes in news framing itself might affect how likely readers are to consume news content: readers might be less likely to consume frames challenging their existing attitudes and reject news outlets containing such frames (and vice versa seek out attitudinally congruent information). This idea of 'selective exposure/avoidance' [@Bennett2008], or a a 'confirmation bias' [@Taber2006] brings with it a major endogeneity problem making purely correlational evidence of framing effects rather uninformative.

Even if we assume that the consumer receives the content, e.g. buys a newspaper, visits its website, or is exposed otherwise, it is not clear whether the content is actually consumed. E.g., a reader might well buy a newspaper but skip an article after reading the headline because the topic is not of interest or the frame is at odds with their existing attitude [@Lau2021a]. As most framing studies do not offer choice of content to respondents, engagement remains an assumption in experimental studies. Additionally, even when consuming the content, motivated reasoners presented with a plethora of news content about an issue will spend less time and cognitive energy on counter-attitudinal arguments than those supporting their viewpoint, lending more importance to considerations raised by amenable content [@Taber2006]. Experimental studies presenting only one frame per issue cannot take this bias into account.

These assumptions of exposure, consumption, and equal engagement in classic framing studies already represent a high hurdle for the translation of experimental estimates of framing effects into the real world. However the differences do not end there. Consumption might result in better or worse comprehension in the real world because respondents might pay more or less attention to real news content compared to experimental treatments [@Barabas2010]. Even after understanding the frame, there might be differences in how easily this translates into changing cognitive associations, given that those exposed might take real-world content more seriously.

To summarize, framing might have very different consequences in real-world environments than their experimental emulations. When we take the experimental evidence as an approximation of the effectiveness of framing in the real world, we make strong assumptions about citizens' exposure to, consumption of, and engagement with the frames in news content. Experiments can only incorporate few characteristics of the real world at a time. In order to test the external validity of framing effect studies, it is necessary to study observational data.

## Existing Observational Studies of Framing Effects

Several scholars have arrived at this conclusion and called for the assessment of framing effects beyond experimental studies [@Busby2019; @Kinder2007; @Leeper2020], yet the call has rarely been answered. While a large body of media effects research exists, it has not paid specific interest to framing, instead assessing the impact of changing news exposure broadly in observational designs [see e.g. @Foos2022; @Grossman2022 @Levendusky2022; @Spirig2020] and field experiments [@Bail2018; @Broockman2022; @Guess2021].

The precious small body of existing research on framing effects is often riddled with endogeneity concerns or compound treatments. For example, @Jerit2008 investigates the effects of framing on public opinion compared to addressing the opponent's issue. She finds that reframing is less effective than engagement with the opposite parties' frames, a finding that suggests that framing is not very effective. However, this study does not provide an estimate of a causal effect, instead restricting itself to the assessment of correlational evidence. This is especially problematic regarding framing effects, as news content can affect consumers' inclination to consume news (see preceding section), and political actors might react to public opinion. Others show with similar correlational designs that only specific frames seem to drive issue importance [@Mclaren2018], but do not assess effects on attitudes.

To my knowledge, there are only two observational studies assessing the causal effect of exposure to specific content on political attitudes. @Slothuus2010b takes advantage of a sudden shift in a political parties' framing of retirement benefits, and shows that this changed framing has substantial impact on party supporters' attitudes about the issue. However, it is unclear how these findings translate to news framing effects, as it is impossible to distinguish the effect of partisan cue-taking from the persuasive effect of the frame itself. The only study which assesses the precise impact of changing news content on political attitudes comes from @Djourelova2020. She exploits a ban of the term "illegal migrant" in the *Associated Press* and shows that this language has modest effects on migration attitudes. While the study itself reveals impressively how language can shape political attitudes, it constitutes a highly specific case and a rather minor change in the framing of migration. While news stories might lack the term "illegal migrant", they might still extensively cover crimes committed by migrants and thus associate migration with crime.

Summarized, we lack an understanding of how framing affects political opinions in the real world. While experimental evidence suggests rather strong effects, it is unclear how these findings translate to the outside world. Observational evidence on media effects is abundant but has not studied precisely what the impact of framing is. The present study therefor exploits a naturally occurring change in the news framing of a particular issue *within* a specific outlet to isolate the effects of news framing on political opinions.

@@

In order to test whether changing news content - rather than a shift in the consumed news outlet - affected political attitudes, I exploit an editorial change in the largest German newspaper, *Bild*. As I will show, this editorial change resulted in a substantial shift in the papers' migration framing and is hence ideal to assess the effect of within-newspaper content shifts. The fielding of a large-scale panel survey in the same period allow to assess the impact of news content on migration attitudes.

@@

In December 2016, Axel Springer, the company owning Bild, announces that Diekmann will leave the company. Less than a week after, it becomes public that prosecutors investigate Diekmann for sexual harassment. While the company claimed that this was not the reason for Diekmann's departure, the process might have been sped up as a result of the investigations\footnote{According to a former editor of the newspaper I talked to, rumors in the company at the time support this interpretation of events.} [@Spiegel2017Diekmann]. The migration coverage in *Bild* severely changed when Diekmann left the newspaper in February 2017 and Julian Reichelt took over the position as editor-in-chief. The paper increasingly started painting a picture of refugees as criminals on the tabloid's front page [@Niggemeyer2018Bild; @DLF2018Reichelt]. Headlines like "I killed 40 people and want asylum" increasingly dominated the migration coverage of the newspaper. In a speech months before he took over the position as chief editor, Reichelt outlined his motivation for such a shift in migration coverage: "I can assure you: nothing has hurt [*Bild*] economically as much as our clear, humane, empathetic stance in the refugee crisis" \footnote{Reichelt in a speech at Deutschlandfunk's conference "Formate des Politischen 2016": \url{https://vimeo.com/190347766}}. He also argued that the German media shows too little tolerance towards anti-immigration attitudes and therefor excludes those holding such views from the political discourse.

This presents a unique case to study the effects of media content on public opinion. The shift in content can be clearly identified. The change of editor itself was not discussed in the media, which means any opinion changes among readers are the result of the changing content [@Spirig2020]. Additionally, the timing of the editorial change is potentially exogenous, given the pressure from allegations against Diekmann. In summary, *this case is as close as possible to a field experiment in which a specific consideration of migration coverage is amplified in the coverage of one newspaper, but not others*.

## Measuring Frames in Communication

To measure immigration framing in the news, I collected 2.5 million newspaper articles from the websites of the most important German broadsheets \textit{Frankfurter Allgemeine Zeitung (FAZ), Spiegel Online (SPON), Süddeutsche Zeitung (SZ), Die Tageszeitung (TAZ), Die Welt} and the major German tabloid \textit{Bild} for the period 2013-2019 (for the market reach of each outlet, see table 1). Following the value-expectancy model, I identify crime frames in migration coverage in a two-step process: first, I use supervised machine learning to identify content about migration. Second, I identify content about crime *within* migration coverage by applying a similar supervised classifier. This approach closely resembles the value-expectancy model, as I first identify coverage about the *issue*, before identifying the considerations emphasized in the discussion of this issue.

To identify whether an article is about migration, i employ supervised machine learning. As migration content is rare, and training data needs to be somewhat balanced, articles cannot be selected randomly for annotation[^1]. Instead, I first construct a dictionary of terms related to migration. I use dictionary extension based on German GloVe word-embeddings\footnote{Downloadable from \url{https://deepset.ai/german-word-embeddings}.} to construct a comprehensive dictionary, and apply it to the articles. Based on the relative share of migration words in an article, I draw a stratified sample of 300 articles (100 from articles with no migration terms, 100 from the quarter with most migration terms and 100 from the remainder in between) for each newspaper for a total of 1,800 papers. Then, a coder hand-coded these articles, assessing whether their main topic is related to migration.

[^1]: If migration content made up about 3% of all articles and the sample were selected randomly, one would find around 30 migration articles per 1000 articles, which provides rather limited information about the class of interest and results in worse recall, as the best guess in highly imbalanced data is simply the most common outcome.

Using this sample, a BERT transformer model\footnote{\url{https://huggingface.co/bert-base-german-cased}} is fine-tuned on a random subset of 1,400 annotated articles (training set) with a validation set of 200 articles. After fine-tuning, the model correctly classifies 95.5% of a test set of 200 articles (F1: 0.94, recall: 0.93, precision: 0.95). This classifier is then used to annotate all 2.5 million newspaper articles. For the period from 2013 to 2019, around 90,000 (3.6%) are identified to treat the issue of migration.

@@

Figure \@ref(fig:salience) shows the share of migration articles in each newspaper. Clear peaks in attention are visible surrounding the refugee movements of summer 2015, as well as following the large-scale sexual assaults on new years eve in Cologne 2016[^2], and a conflict over asylum policy between the conservative sister parties CDU and CSU, which almost lead to the breakdown of the governing coalition in 2018. Newspapers seem to be event-driven and follow largely parallel trends, albeit at different levels, with the centre-left \textit{SZ} discussing migration most and the tabloid \textit{Bild} least. Attention decreases in mid-2016, but remains high compared to pre-2015 levels. Surrounding the fielding period of the survey waves (gray vertical lines), most newspapers show similar levels of attention to the issue between 2.5% and 6% of all articles, with the exception of \textit{SZ}, which moves between 5% and 12.5%.

[^2]: Germany had a large debate about refugees and crime after the new year's eve 2015/16 in Cologne. According to a report of the ministry of interior in Northrhine-Westphalia, hundreds of women were sexually assaulted and robbed by men they perceived to be of "north-african/arabic" descent (see report of the ministry of interior in Northrhine-Westphalia: \\url{[https://web.archive.org/web/20160116104020/http://m.mik.nrw.de/fileadmin/user_upload/Redakteure/Dokumente/Themen_und_Aufgaben/Schutz_und_Sicherheit/160111ssia/160111berppkoeln.pdf}](https://web.archive.org/web/20160116104020/http://m.mik.nrw.de/fileadmin/user_upload/Redakteure/Dokumente/Themen_und_Aufgaben/Schutz_und_Sicherheit/160111ssia/160111berppkoeln.pdf%7D%7D))

To identify migration coverage containing the crime consideration and thereby associating migration with crime, I first fine tune a BERT classifier for German language in a similar fashion as above, this time with coverage about crime as the outcome of interest. Using a training set of 1400 articles, a validation set of 200 and a test set of 200 articles, the classifier reaches a very good performance with an F1 score of 85.7%. I then apply this classifier to classify the approximately 90,000 articles about migration identified by the first classifier. About 7% of all migration coverage consists of articles about some form of crime\footnote{See appendix 1.2 for more information about the classifier, training process and data.}. The share of migration content devoted to news about crime within a given time frame allows me to assess how prominent the consideration of crime as a result of migration was at a specific point in time. In order to give an example of the content, table \@ref(tab:exampleArticles) displays the headlines of the five migration articles within *Bild* most likely discussing crime according to the classifier. All articles refer to crimes of migrant groups, with two referring to violence in refugee homes, two discussing police raids regarding asylum fraud, and one discussing a police report about new years eve in cologne.
